\mythesistitleNOTSYNPOSIS

### 2013-03-22 src/response_to_reviewers.tex

\documentclass{article}
\usepackage[normalem]{ulem}
\usepackage{../lib/mythesislib}
\usepackage{../lib/mylatexlib}
%
\begin{document}

\title{Response to M.S. Thesis Review Comments\\ 
 {Thesis Title: {\em Generalization of the Consecutive-ones Property}}}
\author{Anju Srinivasan\\IIT Madras, Chennai - 600036.}
\date{25 March 2013}

@@

\section{Reviewers}

\begin{enumerate}
\item  {\bf Dr. C. R. Subramaniam}\\ Institute of Mathematical Sciences, Chennai
\item  {\bf Dr. Subhas Chandra Nandy}\\ Indian Statistical Institute, Kolkata
\end{enumerate}

\section{Response to Reviewer 1}

\subsection{Comments on corrections}
Responses to the comments in review document Page 2 items 0 to 7.

\begin{enumerate}
\item[0.] Comment: {\bf In thesis Page 7, Definition 1.3.7: After $x_m
    \le a$, you need to add $x_m \ne a$.}\\
  Response: True, earlier definition did not accomodate the fact $x_m
  \le x_m$. Suggested change made in Page 7 to the definiton of Poset.

@@

\item Comment: {\bf In thesis Page 7, Definition 1.3.8: $E$ is a set
    of ``unordered'' pairs. You need to add the
    word ``unordered''}\\
  Response: Suggested change made in to definition of Undirected Graph
  Page 7 Definition 1.3.8 (1).

\item Comment: {\bf In thesis Page 8, A tree is a {\em connected} and
    acyclic graph}\\
  Response: The word ``connected'' has been added to the definition of
  Tree in Page 7 Definition 1.3.8 (3).

\item Comment: {\bf In thesis Page 8, Definition 1.3.10 you need to
    add that one is looking at intervals of integers. Also define
    isomorphism between graphs for the sake of completeness}.\\
  Response: Definition of Interval graph has been changed to say ``a
  set of intervals (of integers)'' in Page 8, Definition 1.3.10. Graph
  isomorphism definition has been added in Page 7, Definition 1.3.8(4)
  as follows:

  {\em A \textbf{graph isomorphism} between two graphs $G_1$ and $G_2$
    is a bijection $\phi: V(G_1) \rightarrow V(G_2)$ such that $u, v
    \in V(G_1)$ are adjacent in $G_1$ if and only if $\phi(u),
    \phi(v)$ are adjacent in $G_2$. If such a bijection exists, it is
    said that $G_1$ is \textbf{isomorphic} to $G_2$, denoted by $G_1
    \cong G_2$.}

@@

\item Comment: {\bf In thesis Page 8, Definition 1.3.10 definition of
    path graph: You need to emphasise that $\cP$ forms a set system}\\
  Response: This change has been made - ``a set system of paths
  $\cP$'' in Page 8 Definition 1.3.10(2).

\item Comment: {\bf In thesis Page 12, third para of Section 1.5: you
    need to write it more precisely and carefully conveying the
    idea. As it is it is not clear what you are saying.}\\
  Response: The problem {\sc Compute feasible tree path labeling} and
  its result have been described formally and precisely in Chapter 3
  (New results). The aforementioned paragraph describes the same
  informally. These informal sentences have been footnoted with the
  corresponding definitions and theorems from Chapter 3. This change
  is in Page 12.

\item Comment: {\bf In thesis Page 28, Definition 7 (of binary
    operator /): it is not defined precisely. What do you mean by
    ``$b$ is a new element added..''?}.\\
  Response: This was in thesis Page 26. $A/B$ is created by removing
  the elements of second operand set $B$ from the first operand set
  $A$ and introducing a new element {\em not} in the universe to $A$
  to represent set $B$. This is a set operation mentioned in
  \cite{mm96} (they use ``$B$'' to denote the new element to represent
  set $B$ which is also confusing) which is eventually used to
  decompose permutation of elements in $A$ to permutation of elements
  in $A/B$ and permutation of elements in $B$.  Perhaps it would be
  clearer if $b$ is replaced by $x_B$. This change has been made in
  Page 26 Definition 2.2.4 (7) and other mentions of this operation
  ``/'' Page 27 item (iii) and $h$ has been replaced with $x_H$ in
  Corollary 2.2.8. There was also a typo in Page 27 paragraph under
  Corollary 2.2.8 item (ii) where ``$U \setminus H$'' was written
  instead of $U / H$.  This has also been corrected.

\item Comment: {\bf In thesis Page 39 Definition 2.3.1: the definition
    of $A_\cF$ is not clear. Where is the dependence on $\cF$?}\\
  Response: This was in thesis Page 36. It is true, there is no
  dependence in $\cF$. It is only dependent on the universe $U$ of
  $\cF$. All occurences of $A_\cF$ has been changed to $A_U$ in Page
  36.
\end{enumerate}

\subsection{Other comments}
The following are a few observations mentioned in the review document
with my responses/changes.

@@

\begin{enumerate}
\item Observation: (in Page 1, second last paragraph) {\bf The result
    on characterization of TPL in the thesis is a variant of the 1978
    JCT(B) result due to Fournier who showed that two hypergraphs $H_1
    = (V_1, X)$ and $H_2 = V_2, Y$ such that $|X| = |Y| = m$ are
    isomorphic iff there is a bijection $\phi$ on $I = \{1, ... , m\}$
    s.t. for every $L \subset I,
    |\bigcap_{i \in L}X_i| = |\bigcap_{j=\phi(i),i \in L}Y_j|$} \\
  Response: This is an important observation and was not cited in the
  thesis. Also \cite{f80} generalizes \cite{br72} and \cite{fg65} by
  characterizing the isomorphism of two hypergraphs by means of
  equicardinality of certain edge intersections and the exclusion of
  certain pairs of subhypergraphs. TPL characterization is a special
  case of this with one of the hypergraphs having hyperedges that are
  paths from a tree and the characterization only uses edge
  intersections of at most 3 hyperedges. This has been added in the
  Conclusions chapter Page 72 under {\em Graph isomorphism and
    logspace canonization}.

\item Observation: (in Page 2, first paragraph) {\bf A $k$-subdivided
    tree is a tree with a single central vertex with a number of paths
    each of $k$ edges emanating from it.}\\
  Response: $k$-subdivided as defined in this thesis (Section 3.2.4,
  Page 43) is such that each of the aforementioned paths are of length
  ``$k+1$'' edges since each such path has $k$ nodes apart from the
  central node and the leaf node, i.e.  $k+2$ nodes. No change in the
  thesis.
\end{enumerate}

\section{Response to Reviewer 2}

\section{Other comments}
\begin{enumerate}
\item Observation: {\bf The definition of $k$-subdivided star is not
    given in the thesis. What I understood is that it has a root node $v$
    of degree $k$, the degree of all other nodes is at most 2, and
    each path from root to leaf is $k+2$.}\\
Response: The definition of $k$-subdevided star was given in Section
3.2.4, Definition 3.2.6. Perhaps this has not been sufficiently
referred to, hence references to this definition has been made in
Chapter 1 Page 13 second last paragraph where it is first referred and
in Chapter 3 Page 39 second paragraph where {\sc Compute
  $k$-subdivided Star Path Labeling} problem is defined.

The reviewer's understanding is 
correct except that the root node can have any degree, thus a
$k$-subdivided star is a collection of graphs with degree of all
non-root nodes as at most 2, each path from root to leaf is $k+2$ and 
the root having any degree. See thesis Figure 1.5 in page 13 for examples.

\item Comment: {\bf The proof of Lemma 3.3.4 is unclear.}\\
Response: The proof was correct but terse. It has been elaborated to as
follows:

  {\em 
  Lemma:   Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
  \leq i \leq 4$.  Then, $|\cap_{i=1}^4 S_i| = |\cap_{i=1}^4 P_i|$.

  Proof: Consider the set of set intersections with $S_1$ $\cF' = \{S_2 \cap S_1, S_3
  \cap S_1, S_4 \cap S_1\}$ and let $\cl'$ be a tree path labeling of
  $\cF'$ such that $\cl'(S_i \cap S_1) = P_i \cap P_1, 2 \le i \le
  4$. Clearly, $P_2 \cap P_1$, $P_3 \cap P_1$, and $P_4 \cap P_1$ are
  subpaths of path $P_1$, thus equivalent to intervals.  Due to the
  three way intersection cardinality preservation property of the
  ICPPL $(\cF, \cl)$, this new tree path labeling $(\cF', \cl')$
  preserves pairwise intersection cardinalities.  Now by applying
  Lemma 3.3.1 %~\ref{lem:intersection-cardinality} 
  to the sets in $\cF'$ and
  their corresponding path images due to $\cl'$, it follows that
  $|\cap_{i=1}^4 S_i| =| \cap_{i=1}^4 P_i|$.}

\item Comment: {\bf While studying Feasible Tree Path Labeling Problem
  in Section 3.3, two filters are mentioned and their pseudocodes
  given. If possible, state the problem in words. ... Explicitly state
the roles of these two filters in the algorithm with proper
figures.}\\
Response:

\item Comment: {\bf It is shown that TREE PATH LABELING PROBLEM on
    arbitrary trees can be solved polynomial time if the two
    subproblems, identified in section 3.6 in the thesis are
    polynomial time solvable. The formal definition of these two
    subproblems are not given. However, in page 71 item numbers 1 and
    2 the subproblems are described with several notations, where the
    motivation is unclear.}\\
Response:

\item Comment: {\bf ... it is shown that consecutive 1s property
    testing for a \sout {graph} [binary matrix] is solvable in LOGSPACE. {\em This is a very
      important contribution.}}\\
Response: Since this is a conclusion from earlier results, it has not
been mentioned as a ``new'' result. It is mentioned in the abstract as a
derived result.
\end{enumerate}

### 2013-03-22 src/010chap1.tex
Most of reviewer 2 comments done.

We present a polynomial time algorithm for {\CFTPL} when the target
tree $T$ belongs to a special class of trees called {\em \kstars} and
when the hyperedges in the hypergraph $\cF$ have at most $k+2$
vertices (\CFTPLKTREE -- see Section~\ref{sec:myresearchintro} for
problem definition). A couple of examples of {\kstars} are given in
Figure~\ref{fig:ksubstar}.

>>>>

We present a polynomial time algorithm for {\CFTPL} when the target
tree $T$ belongs to a special class of trees called {\em \kstars} (see
Definition~\ref{def:ksubstar} for the formal definition) and
when the hyperedges in the hypergraph $\cF$ have at most $k+2$
vertices (\CFTPLKTREE -- see Section~\ref{sec:myresearchintro} for
problem definition). A couple of examples of {\kstars} are given in
Figure~\ref{fig:ksubstar}.

### 2013-03-22 src/030chap3.tex
Most of reviewer 2 comments done.

With respect to computing a feasible TPL, as suggested by \CFTPL
problem, we were unable to discover an efficient algorithm for
it. However, we consider two special cases of the same on special
target trees -- namely, \CFTPLINT and \CFTPLKTREE on intervals and
\kstars, respectively.  Section~\ref{sec:spltargettree} discusses
these problems and their polynomial time solutions.  \CFTPLINT is the
known problem of \COP testing. The polynomial time solution to
\CFTPLKTREE is a new result by us.

@@

\begin{lemma}
  \label{lem:4-way}
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
  \leq i \leq 4$.  Then, $|\cap_{i=1}^4 S_i| = |\cap_{i=1}^4 P_i|$.
\end{lemma}
\begin{proof}
  Consider the sets $S_2 \cap S_1$, $S_3 \cap S_1$, and $S_4 \cap
  S_1$, and their images $P_2 \cap P_1$, $P_3 \cap P_1$, and $P_4 \cap
  P_1$ respectively.  It is clear that this path labeling preserves
  pairwise intersection cardinalities.  In this case, it is clear that
  the 3 sets are assigned paths in the path $P_1$.  Now by applying
  Lemma~\ref{lem:intersection-cardinality}, it follows that
  $|\cap_{i=1}^4 S_i| =| \cap_{i=1}^4 P_i|$.
\end{proof}
\begin{corollary} 
  \label{corr:4-way} 
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
  \leq i \leq 4$. If $P_i$, $1
  \leq i \leq 4$ are such that $P_1 \cap P_2$, $P_1
  \setminus P_2$, $P_2 \setminus P_1$ are also paths, then:
  \begin{align*}
   |(S_1 \setminus S_2) \cap S_3 \cap S_4| &= |(P_1 \setminus P_2) \cap P_3 \cap P_4|\\
   |(S_2 \setminus S_1) \cap S_3 \cap S_4| &= |(P_2 \setminus P_1) \cap P_3 \cap P_4| 
  \end{align*}
\end{corollary}
\begin{proof}
  It is clear that $|(P_1 \setminus P_2) \cap P_3 \cap P_4| =
  |\displaystyle \cap_{i=1}^4 P_i| - | P_2 \cap P_3 \cap P_4|$.  Since
  we have an ICPPL, from Lemma~\ref{lem:4-way}, it follows that the
  intersection cardinalities are preserved.
\end{proof}
\begin{lemma}
  \label{lem:priv-card}
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S) = P$ for some $S \in \cF$.  If sets
  $S_{priv}$ and $P_{priv}$ contain only those elements of $S$ and $P$
  that occur in no set other than $S$ and $P$, respectively, then
  $|S_{priv}|=|P_{priv}|$.
\end{lemma}
\begin{proof}
  Consider the set system $ \cF' = \{S_i \cap S \mid S_i \in \cF, i
  \in [n]\}$ and path labeling $\cl'(S_i \cap S) = P_i \cap P$.  Since
  $(\cF, \cl)$ is an \ICPPL, it can be easily verified that $(\cF',
  \cl')$ is an ICPIA on path (interval) $P$.  Let $S_{two}$ and
  $P_{two}$ be the elements of $S$ and $P$ which are in one {\em
    other} set and path, respectively \ie $S_{two} = S \cap (\cup_{S_i
    \ne S} S_i)$ and $P_{two}$ is defined analogously. It may be noted
  that $S_{two} = \supt{\cF'}$ and $P_{two} = \supt{\cF'^{\cl'}}$.
  From Lemma~\ref{lem:icpia} it follows that there is a bijection
  $\phi$ from $S_{two}$ to $P_{two}$ such that for each $i$, the image
  of $S_i \cap S$ under $\phi$ is $P_i \cap P$ (definition of
  feasibility).  Moreover, this means $|S_{two}| = |P_{two}|$. By
  definition, $S_{priv} = S \setminus S_{two}$ and $P_{priv} = P \setminus P_{two}$.
  Thus $|S_{priv}| = |S \setminus S_{two}| = |P \setminus
  P_{two}|=|P_{priv}|$.  
\end{proof}

@@

In this section we consider the problem of assigning paths from a
$k$-subdivided star $T$ to a given set system $\cF$ as described by
the \CFTPLKTREE problem.  We consider
$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
overlap graph is well-known from the work of
\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
and $S'$ have a non-empty intersection but neither of them is
contained in the other. The overlap graph $\bO(\cF)$ is a graph in
which the vertices correspond to the sets in $\cF$, and the vertices
corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
if they overlap.  Note that the intersection graph of $\cF$,
$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
component of $\cF$.  An interesting property of the overlap components
is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
one set in $\cO_2$.  This containment relation naturally determines a
decomposition of the overlap components into rooted containment trees.
We consider the case when there is only one rooted containment tree,
and we first present our algorithm when $\bO(\cF)$ is connected.  It
is easy to see that once the path labeling to the overlap component in
the root of the containment tree is achieved, the path labeling to the
other overlap components in the rooted containment tree is essentially
finding a path labeling when the target tree is a path: each target
path is a path that is allocated to sets in the root overlap
component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
connected graph. We also assume that all hyperedges are of cardinality
at most $k+2$.

>>>>

With respect to computing a feasible TPL, as suggested by \CFTPL
problem, we were unable to discover an efficient algorithm for
it. However, we consider two special cases of the same on special
target trees -- namely, \CFTPLINT and \CFTPLKTREE on intervals and
\kstars (see Definition~\ref{def:ksubstar}), respectively.  Section~\ref{sec:spltargettree} discusses
these problems and their polynomial time solutions.  \CFTPLINT is the
known problem of \COP testing. The polynomial time solution to
\CFTPLKTREE is a new result by us.

@@

\begin{lemma}
  \label{lem:4-way}
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
  \leq i \leq 4$.  Then, $|\cap_{i=1}^4 S_i| = |\cap_{i=1}^4 P_i|$.
\end{lemma}
\begin{proof}
  Consider the set of set intersections with $S_1$ $\cF' = \{S_2 \cap S_1, S_3
  \cap S_1, S_4 \cap S_1\}$ and let $\cl'$ be a tree path labeling of
  $\cF'$ such that $\cl'(S_i \cap S_1) = P_i \cap P_1, 2 \le i \le
  4$. Clearly, $P_2 \cap P_1$, $P_3 \cap P_1$, and $P_4 \cap P_1$ are
  subpaths of path $P_1$, thus equivalent to intervals.  Due to the
  three way intersection cardinality preservation property of the
  ICPPL $(\cF, \cl)$, this new tree path labeling $(\cF', \cl')$
  preserves pairwise intersection cardinalities.  Now by applying
  Lemma~\ref{lem:intersection-cardinality} to the sets in $\cF'$ and
  their corresponding path images due to $\cl'$, it follows that
  $|\cap_{i=1}^4 S_i| =| \cap_{i=1}^4 P_i|$.
\end{proof}
\begin{corollary} 
  \label{corr:4-way} 
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S_i)=P_i$, $S_i \in \cF$, $1
  \leq i \leq 4$. If $P_i$, $1
  \leq i \leq 4$ are such that $P_1 \cap P_2$, $P_1
  \setminus P_2$, $P_2 \setminus P_1$ are also paths, then:
  \begin{align*}
   |(S_1 \setminus S_2) \cap S_3 \cap S_4| &= |(P_1 \setminus P_2) \cap P_3 \cap P_4|\\
   |(S_2 \setminus S_1) \cap S_3 \cap S_4| &= |(P_2 \setminus P_1) \cap P_3 \cap P_4| 
  \end{align*}
\end{corollary}
\begin{proof}
  It is clear that $|(P_1 \setminus P_2) \cap P_3 \cap P_4| =
  |\displaystyle \cap_{i=1}^4 P_i| - | P_2 \cap P_3 \cap P_4|$.  Since
  we have an ICPPL, from Lemma~\ref{lem:4-way}, it follows that the
  intersection cardinalities are preserved.
\end{proof}
\begin{lemma}
  \label{lem:priv-card}
  Let $(\cF, \cl)$ be an \ICPPL and $\cl(S) = P$ for some $S \in \cF$.  If sets
  $S_{priv}$ and $P_{priv}$ contain only those elements of $S$ and $P$
  that occur in no set other than $S$ and $P$, respectively, then
  $|S_{priv}|=|P_{priv}|$.
\end{lemma}
\begin{proof}
  Consider the set system $ \cF' = \{S_i \cap S \mid S_i \in \cF, i
  \in [n]\}$ and path labeling $\cl'(S_i \cap S) = P_i \cap P$.  Since
  $(\cF, \cl)$ is an \ICPPL, it can be easily verified that $(\cF',
  \cl')$ is an ICPIA on path (interval) $P$.  Let $S_{two}$ and
  $P_{two}$ be the elements of $S$ and $P$ which are in one {\em
    other} set and path, respectively \ie $S_{two} = S \cap (\cup_{S_i
    \ne S} S_i)$ and $P_{two}$ is defined analogously. It may be noted
  that $S_{two} = \supt{\cF'}$ and $P_{two} = \supt{\cF'^{\cl'}}$.
  From Lemma~\ref{lem:icpia} it follows that there is a bijection
  $\phi$ from $S_{two}$ to $P_{two}$ such that for each $i$, the image
  of $S_i \cap S$ under $\phi$ is $P_i \cap P$ (definition of
  feasibility).  Moreover, this means $|S_{two}| = |P_{two}|$. By
  definition, $S_{priv} = S \setminus S_{two}$ and $P_{priv} = P \setminus P_{two}$.
  Thus $|S_{priv}| = |S \setminus S_{two}| = |P \setminus
  P_{two}|=|P_{priv}|$.  
\end{proof}

@@

In this section we consider the problem of assigning paths from a
$k$-subdivided star (see Definition~\ref{def:ksubstar}) $T$ to a given
set system $\cF$ as described by the \CFTPLKTREE problem.  We consider
$\cF$ for which the overlap graph $\bO(\cF)$ is connected.  The
overlap graph is well-known from the work of
\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
\cite{kklv10}. Recall from Section~\ref{ch:prelims} that hyperedges
$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$
and $S'$ have a non-empty intersection but neither of them is
contained in the other. The overlap graph $\bO(\cF)$ is a graph in
which the vertices correspond to the sets in $\cF$, and the vertices
corresponding to the hyperedges $S$ and $S'$ are adjacent if and only
if they overlap.  Note that the intersection graph of $\cF$,
$\bI(\cF)$ is different from $\bO(\cF)$ and $\bO(\cF) \subseteq
\bI(\cF)$.  A connected component of $\bO(\cF)$ is called an overlap
component of $\cF$.  An interesting property of the overlap components
is that any two distinct overlap components, say $\cO_1$ and $\cO_2$,
are such that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are
disjoint, or, w.l.o.g, all the sets in $\cO_1$ are contained within
one set in $\cO_2$.  This containment relation naturally determines a
decomposition of the overlap components into rooted containment trees.
We consider the case when there is only one rooted containment tree,
and we first present our algorithm when $\bO(\cF)$ is connected.  It
is easy to see that once the path labeling to the overlap component in
the root of the containment tree is achieved, the path labeling to the
other overlap components in the rooted containment tree is essentially
finding a path labeling when the target tree is a path: each target
path is a path that is allocated to sets in the root overlap
component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
connected graph. We also assume that all hyperedges are of cardinality
at most $k+2$.
